---
title: "Analysis Report"
author: "Svenja Kuster"
date: today
toc: true
abstract-title: "Research questions:"
abstract: "Does the new drug AH-27 lead to a greater reduction of systolic blood pressure (SBP) compared to standard care? Is a higher baseline SBP associated with a greater reduction in SBP? Do older patients have a reduced responsiveness to AH-27? Is there a sex difference in the response to AH-27?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

# Prepare Session and Import/Clean Data

```{r}
#| label: load-packages
#| echo: false
#| output: false

library(pastecs)
library(ggpubr); theme_set(theme_bw())
library(GGally)
library(performance)
library(MuMIn)
library(ggeffects)
library(knitr)
```

```{r}
#| label: import-and-clean-data
#| echo: false
#| output: false

sbp <- read.csv("./data/processed/hyper-tension.csv", stringsAsFactors = T)
str(sbp)
```

# Explore the Data

## numerically

```{r}
#| label: tbl-numerical-exploration
#| tbl-cap: "Descriptive statistics of variables of interest"
#| echo: false

desc.stats <- round(stat.desc(sbp, norm=T)[c("reduction_sbp", "age", "baseline_sbp")], 2)
# table(sbp$sex, sbp$treatment_group) #design seems balanced

kable(desc.stats) #data looks good and assumption of normality seems fullfilled, except for age, bc the age of people who get treated for high blood pressure is obviously very similar
```

As shown in @tbl-numerical-exploration, the assumption of normality is fulfilled for the two blood pressure variables but not for "age". This, however, is expected, since the age of patients who get treated for high blood pressure, will be very similar (i.e. very young patients for example will usually not be treated for hypertension).

## graphically

```{r}
#| label: graphical-exploration
#| echo: false
#| output: false

ggpairs(sbp[,-1])
```

```{r}
#| label: fig-main-research-question
#| fig-cap: "Visualisation of main research question including possible confounder variable sex"
#| fig-width: !expr 12/2.54
#| fig-height: !expr 9/2.54
#| echo: false

ggplot(sbp, aes(treatment_group, reduction_sbp, col=sex))+
  geom_jitter(width = .2, alpha=.2)+
  geom_boxplot(alpha=0)+
  labs(x="Treatment Group", y="Reduction in Systolic Blood Pressure [mmHg]", title = "Reduction in Systolic Blood Pressure According to Treatment Group and Sex")
```

# Fit Candidate Models and Select the Best One

# Validate Preferred Model

# Interpret and Report Findings

# Store Session Information
