---
title: "Analysis Report"
author: "Svenja Kuster"
date: today
toc: true
abstract-title: "Research questions:"
abstract: "Does the new drug AH-27 lead to a greater reduction of systolic blood pressure (SBP) compared to standard care? Is a higher baseline SBP associated with a greater reduction in SBP? Do older patients have a reduced responsiveness to AH-27? Is there a sex difference in the response to AH-27?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

# Prepare Session and Import/Clean Data

```{r}
#| label: load-packages
#| echo: false
#| output: false

library(pastecs)
library(ggpubr); theme_set(theme_bw())
library(GGally)
library(performance)
library(MuMIn)
library(ggeffects)
library(knitr)
```

```{r}
#| label: import-and-clean-data
#| echo: false
#| output: false

sbp <- read.csv("./data/processed/hyper-tension.csv", stringsAsFactors = T)
sbp$treatment_group <- factor(sbp$treatment_group, levels=c("standard_care", "AH27")) #change order of factors for easier interpretation later on
str(sbp)
```

# Explore the Data

## Numerically

```{r}
#| label: tbl-numerical-exploration
#| tbl-cap: "Descriptive statistics of variables of interest"
#| echo: false

desc.stats <- round(stat.desc(sbp, norm=T)[c("reduction_sbp", "age", "baseline_sbp")], 2)
# table(sbp$sex, sbp$treatment_group) #design seems balanced

kable(desc.stats) #data looks good and assumption of normality seems fullfilled, except for age, bc the age of people who get treated for high blood pressure is obviously very similar
```

As shown in @tbl-numerical-exploration, the assumption of normality is fulfilled for the two blood pressure variables but not for "age". This, however, is expected, since the age of patients who get treated for high blood pressure, will be very similar (i.e. very young patients for example will usually not be treated for hypertension).

## Graphically

```{r}
#| label: graphical-exploration
#| echo: false
#| output: false

ggpairs(sbp[,-1])
```

```{r}
#| label: fig-main-research-question
#| fig-cap: "Visualisation of main research question including possible confounder variable sex"
#| fig-width: !expr 12/2.54
#| fig-height: !expr 9/2.54
#| echo: false

ggplot(sbp, aes(treatment_group, reduction_sbp, col=sex))+
  geom_jitter(width = .2, alpha=.2)+
  geom_boxplot(alpha=0)+
  labs(x="Treatment Group", y="Reduction in Systolic\nBlood Pressure [mmHg]", title = "Reduction in Systolic Blood Pressure\nAccording to Treatment Group and Sex")
```

@fig-main-research-question depicts the main research question of this study and also includes a possible confounding variable, which is sex. This figure already shows us, that the advantage of the new drug AH-27 over the current gold standards is probably negligible.

# Fit Candidate Models and Select the Best One

```{r}
#| label: fit-candidate-models
#| output: false

mod.null <- lm(reduction_sbp~1, data=sbp) #null model for comparison
mod.additive <- lm(reduction_sbp~age+baseline_sbp+treatment_group+sex, data=sbp) #all variables of interest additive
mod.int.age <- lm(reduction_sbp~baseline_sbp+treatment_group*age +sex, data=sbp) #interaction between treatment group and age
mod.int.sex <- lm(reduction_sbp~baseline_sbp+treatment_group*sex +age, data=sbp) #interaction between treatment group and sex
mod.int.sex.age <- lm(reduction_sbp~baseline_sbp+treatment_group*sex *age, data=sbp) #interaction between treatment group, sex and age
```

```{r}
#| label: model-comparison
#| output: false

mod.comparison <- compare_performance(mod.null, mod.additive, mod.int.age, mod.int.sex, mod.int.sex.age)

mod.final <- mod.int.age
```

The AICc (`r round(mod.comparison[2,5], 2)`) and BIC (`r round(mod.comparison[2,7], 2)`) is the best for the additive model. When looking at the adjusted *R*^2^ , the model that includes an interaction term with age (*R*^2^ = `r round(mod.comparison[3,10], 3)`), has a marginally higher *R*^2^ than the additive model (*R*^2^ = `r round(mod.comparison[2,10], 3)`). However, since the interactive effect of age and treatment is included in our research question, I will be continuing using the model that includes this interaction term.

# Validate Preferred Model

```{r}
#| label: fig-model-validation
#| fig-cap: "Graphical evaluation of model performance"
#| fig-width: !expr 20/2.54
#| fig-height: !expr 24/2.54

check_model(mod.final)
```

@fig-model-validation shows that overall, our model performs quite well and fulfills our assumptions. Note that the high VIF is to be expected, since we included an interaction term in our model.

# Interpret and Report Findings

```{r}
#| label: model-interpretation
#| echo: false

#coefficients table:
coef.table<- round(summary(mod.final)$coefficients, 3)

#r squared:
r2 <- round(summary(mod.final)$r.squared,3)

#comparison to null model:
null.chi <- anova(mod.null, mod.final)

#plot that shows effect of baseline sbp on reduction:
pred.baseline <- predict_response(mod.final)$baseline_sbp
plot.baseline <- ggplot(data=pred.baseline, aes(x, predicted))+
  geom_line(col="darkgreen")+
  geom_ribbon(aes(ymin= conf.low, ymax= conf.high), alpha=.3, fill="darkgreen")+
  geom_point(data=sbp, aes(x=baseline_sbp, y=reduction_sbp), alpha=.2)+
  labs(x="Baseline Systolic Blood Pressure [mmHg]", y="Reduction in Systolic Blood Pressure [mmHg]", title="a) Effect of Baseline Systolic Blood Pressure\non the Reduction of Systolic Blood Pressure")

#plot that shows the effect of age*treatment group on reduction:
p <- seq(min(sbp$age), max(sbp$age), length.out= 100)
pred.age.treatment <- predict_response(mod.final, c("age", "treatment_group"))
plot.age.treatment <- ggplot(data=pred.age.treatment, aes(x, predicted))+
  geom_ribbon(aes(ymin= conf.low, ymax= conf.high, fill= group), alpha= .2) +
    geom_line(aes(col= group))+
  geom_point(data=sbp, aes(x=age, y=reduction_sbp, col=treatment_group), alpha=.2)+
  labs(x="Age", y="Reduction in Systolic Blood Pressure [mmHg]", title="b) Effect of Age and Treatment\non the Reduction of Systolic Blood Pressure",col="Treatment", fill="Treatment")
```

```{r}
#| label: tbl-model-interpretation
#| tbl-cap: !expr paste("Estimates, associated standard errors and statistical significance of the parameters estimated by a linear model.", "($R^2$=", r2, ", $\\chi^2_{(", null.chi[2, 3], ",", null.chi[2, 1], ")}$=", round(null.chi[2,5], 3), ", $p$", format.pval(null.chi[2, 6], eps = 0.001), ")")
#| echo: false

kable(coef.table)
```

```{r}
#| label: fig-report
#| echo: false
#| fig-width: !expr 24/2.54
#| fig-height: !expr 12/2.54
#| fig-cap: "Effect of Baseline Systolic Blood Pressure, Age and Treatment on the Reduction of Systolic Blood Pressure"

ggarrange(plot.baseline, plot.age.treatment)
```

A linear model (*R*^2^ = `r r2` , `r paste("$\\chi^2_{(", null.chi[2,3], ",", null.chi[2,1], ")}$")`= `r round(null.chi[2, 5], 3)` , *p*`r format.pval(null.chi[2, 6], eps= 0.001)`; @tbl-model-interpretation) performed significantly better than a null model, comprising an intercept only. The model showed a significant effect of only baseline systolic blood pressure ($\beta_{baseline-sbp}$= `r round(coef.table[2, 1], 3)` , *SE*= `r round(coef.table[2, 2], 3)` , *z*= `r round(coef.table[2, 3], 3)`, *p*= `r round(coef.table[2,  4], 3)`; @fig-report *a)*) on the reduction of systolic blood pressure.

However, the model generated no compelling support for the treatment ($\beta_{treatment-groupAH27}$= `r round(coef.table[3, 1], 3)` , *SE*= `r round(coef.table[3, 2], 3)` , *z*= `r round(coef.table[3, 3], 3)`, *p*= `r round(coef.table[3,  4], 3)`), age ($\beta_{age}$= `r round(coef.table[4, 1], 3)` , *SE*= `r round(coef.table[4, 2], 3)` , *z*= `r round(coef.table[4, 3], 3)`, *p*= `r round(coef.table[4,  4], 3)`) or sex ($\beta_{sexmale}$= `r round(coef.table[5, 1], 3)` , *SE*= `r round(coef.table[5, 2], 3)` , *z*= `r round(coef.table[5, 3], 3)`, *p*= `r round(coef.table[5,  4], 3)`). Therefore, there is also no interactive effect between age and treatment. These findings are can also be seen in @fig-report *b)*, where there is no noticeable slope for age and no significant change in slope for the two treatment groups.

Overall, as already seen in the first visualisation of our research question (@fig-main-research-question), the new drug AH-27 shows no significant advantage over the current gold standard treatment.

# Store Session Information

```{r}
#| label: session-information
#| echo: false
#| output: false

version <- sessionInfo() 
```
